---
title: "Reproducible Research Project 1"
author: "Jaime Jordan"
date: "August 27, 2016"
output: html_document
---
#Reproducible Research Project 1
##Jaime Jordan

Loading and preprocessing the data:
```{r}
library(dplyr)
```

Attaching package: 'dplyr'

```{r}
setwd("C:/Users/Jaime/Desktop/Coursera/ReproducibleResearch")
activitydata<- read.csv("activity.csv")
head(activitydata)
```

```{r}
dim(activitydata)
```

```{r}
glimpse(activitydata)
```

```{r}
summary(activitydata)
```

```{r}
activitydata$date<- as.Date(activitydata$date) ##convert chr to date
```

##What is the mean total number of steps taken per day?
```{r}
###Calculate the total number of steps taken per day:
Total_Steps<-activitydata%>%
  group_by(date)%>%
  filter(!is.na(steps))%>%
  summarize(total_steps = sum(steps, na.rm=TRUE))
Total_Steps
```

```{r}
##Plot using ggplot
library(ggplot2)
ggplot(Total_Steps, aes(x=total_steps)) +
  geom_histogram(fill = "blue", binwidth = 1000) +
  labs(title = "Daily Steps", x= "Total Steps", y = "Frequency")
```

```{r}
##Calculate and report the mean and median of the total number of steps taken per day
Mean_Steps<- mean(Total_Steps$total_steps, na.rm=TRUE)
Mean_Steps
```

```{r}
Median_Steps<- median(Total_Steps$total_steps, na.rm=TRUE)
Median_Steps
```

##What is the average daily activity pattern?
```{r}
## Calculate the average steps
Interval<- activitydata%>%
  group_by(interval)%>%
  filter(!is.na(steps))%>%
  summarize(avg_steps = mean(steps, na.rm=TRUE))
Interval
```

```{r}
##Plot average steps
ggplot(Interval, aes(x=interval , y=avg_steps)) +
  geom_line(color="blue", size=1) +
  labs(title = "Average Daily Steps", x = "Interval", y = "Average Steps per day")
```

```{r}
###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
Interval[which.max(Interval$avg_steps),]
```

##Imputing Missing Values
```{r}
##Calculate total number of missing values in the dataset
sum(is.na(activitydata$steps))
```

```{r}
##Imputing missing values using mean for each day and create a new dataset that is equal to the original dataset but with the missing data filled in.
activitydata2<- activitydata
nas<- is.na(activitydata2$steps)
avg_interval<- tapply(activitydata2$steps, activitydata2$interval, mean, na.rm=TRUE, simplify = TRUE)
activitydata2$steps[nas] <- avg_interval[as.character(activitydata2$interval[nas])]
names(activitydata2)
```

````{r}
##Check if no missing value is appearing
sum(is.na(activitydata2))
```

```{r}
##Reorder columns (for better understanding of the data)
activitydata2 <- activitydata2[, c("date", "interval", "steps")]
head(activitydata2)
```

##Create histogram of the total number of steps taken each day and calculate mean and median total number of steps taken per day.

```{r}
Total_Steps2<- activitydata2%>%
  group_by(date)%>%
  summarize(total_steps = sum(steps, na.rm=TRUE))
Total_Steps2
```

```{r}
ggplot(Total_Steps2, aes(x = total_steps)) +
  geom_histogram(fill = "blue", binwidth = 1000) +
  labs(title = "Daily Steps Including Missing Values", x = "Interval", y = "No. of Steps")
```

```{r}
##Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
Mean_Steps2<- mean(Total_Steps2$total_steps, na.rm=TRUE)
Mean_Steps2
```

```{r}
Median_Steps2<- median(Total_Steps2$total_steps, na.rm=TRUE)
Median_Steps2
```

```{r}
##Answer: The impact of imputing missing data with the average number of steps in the same 5-minute interval is that both the mean and the median are the same: 10766.19
```

##Are there differences in activity patterns between weekdays and weekends?
```{r}
head(activitydata2)
```

```{r}
##Create new variable called WeekType for Weekday and Weekend
activitydata2<- activitydata2%>%
  mutate(weektype= ifelse(weekdays(activitydata2$date)=="Saturday" | weekdays(activitydata2$date)=="Sunday", "Weekend", "Weekday"))

head(activitydata2)
```

```{r}
##Plotting
Interval2<-activitydata2%>%
  group_by(interval, weektype)%>%
  summarize(avg_steps2 = mean(steps, na.rm=TRUE))
head(Interval2)
```

```{r}
plot<- ggplot(Interval2, aes(x = interval , y=avg_steps2, color=weektype)) +
  geom_line() +
  labs(title = "Average Daily Steps by Weektype", x = "Interval", y = "No. of Steps") +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(plot)
```

```{r}
##Answer: During the weekday, the test object is more active earlier in the day but the object is more active throughout the weekends. This is most likely due to the object working during the weekdays.
```

